Dependency in Refinement Logics

### 2015-10-25 doc/paper.tex
Start taking advantage of ABT framework

@@

The meaning of $\Refine{\JJ}{\tau}{\Psi}{E}$ is that $\tau$ is applicable to
demonstrating the judgment $\JJ$, producing synthesis $e$ under the assumptions
that the judgments in $\Psi$ can be demonstrated. The divergence judgment
$\NoRefine{\JJ}{\tau}$ expresses the inapplicability of $\tau$ to $\JJ$.  In
practice, we will explain only the assertion conditions for one of
$\Refine{\JJ}{\tau}{\Psi}{E}$ and $\NoRefine{\JJ}{\tau}$, and implicitly take
the other to be its complement.

@@

Let $\JSet$ be the open-ended collection of judgments in our logical theory,
and let $\RSet$ be a collection of rule names. Each rule $\Member{R}{\RSet}$
must be interpretable as a tactic, i.e.\ the meaning of the assertions
$\Refine{\JJ}{R}{\Psi}{E}$ and $\NoRefine{\JJ}{R}$ must be explained for
$\Member{J}{\JSet}$. We say \framebox{$\IsTac{\tau}$} in case for all object-judgments
$\Member{\JJ}{\JSet}$, the assertion conditions for
$\Refine{\JJ}{\tau}{\Psi}{E}$ and $\NoRefine{\JJ}{\tau}$ are disjoint, and
moreover, if $\Refine{\JJ}{\tau}{\Psi}{E}$, then
$\SubsetEq{\FV{E}}{\Dom{\Psi}}$.

@@

\begin{thm}
  The above rules all define valid tactics:.
  \begin{enumerate}
    \item $\IsTac{\IdTac}$.
    \item $\IsTac{\FailTac}$.
    \item If $\IsTac{\tau_1}$ and $\IsTac{\tau_2}$, then
      $\IsTac{\OrElseTac{\tau_1}{\tau_2}}$.

    \item If $\IsTac{\tau_1}$ and for any $\Member{\alpha}{I}$,
      $\IsTac{\tau_\alpha}$, then
      $\IsTac{\ThenTac{\tau_1}{\MkFam{\tau}{\alpha}{I}}}$.

     %% If $\IsTac{\tau_1}$ and $\IsTac{\tau_2}$, then
     %% $\IsTac{\ThenTac{\tau_1}{\tau_2}}$.
  \end{enumerate}
\end{thm}
\begin{proof}
  It suffices to verify that the synthesis $E$ of each tactical is well-scoped
  in $\Psi$; in each case, this follows by induction.
  \begin{enumerate}
    \item[(1--2)] Immediate.
    \item[(3)]
      By induction on derivations of
      $\Refine{\JJ}{\OrElseTac{\tau_1}{\tau_1}}{E}{\Psi}$.
      \begin{description}
        \item[Case $\Refine{\JJ}{\tau_1}{\Psi}{E}$.] Validity follows from the
          inductive hypothesis $\IsTac{\tau_1}$.
        \item[Case $\NoRefine{\JJ}{\tau_1}$.] Validity follows from the
          inductive hypothesis $\IsTac{\tau_2}$.
      \end{description}

    \item[(4)] It suffices to verify the cases where
      $\IsEquiv{\Dom\Phi}{I}$, since otherwise
      $\NoRefine{\JJ}{\ThenTac{\tau_1}{\MkFam{\tau}{\alpha}{I}}}$.
      We need to show that
      $\SubsetEq{\FV{\Subst{F}{\Dom\Phi}{E}}}{\Dom{\bigoplus_{\Dom\Phi}\Psi}}$.
      From our inductive hypotheses, it is evident that
      $\SubsetEq{\FV{E}}{\Dom\Phi}$ and that for each $\Member{\alpha}{\Phi}$,
      $\SubsetEq{\FV{F_\alpha}}{\Dom{\Psi_\alpha}}$; all of the free variables
      of the term got by substituting each $F_\alpha$ for
      $\Member{\alpha}{\Dom\Phi}$ clearly reside in one of the fibres of the
      family of contexts $\Psi$, and so they must comprise a subset of the
      union $\bigoplus_{\Dom\Phi}\Psi$ of $\Psi$'s fibres.

@@

>>>>

\newcommand\IsAbt[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \IMode{#3}\vdash%
  \IMode{#4}:\OMode{#5}%
}
\newcommand\IsWf[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \IMode{#3}\vdash%
  \IMode{#4}\in_{\mathbf{wf}}\IMode{#5}%
}

\newcommand\SortJdg{\mathsf{jdg}}
\newcommand\MkBTm[3]{\{#1\}[#2].\, #3}
\newcommand\JdgValence[1]{\mathfrak{v}(#1)}
\newcommand\Metavars[1]{\mathfrak{M}(#1)}

@@

The meaning of $\Refine{\JJ}{\tau}{\Psi}{E}$ is that $\tau$ is applicable to
demonstrating the judgment $\JJ$, producing synthesis $E$ under the assumptions
that the judgments in $\Psi$ can be demonstrated. The divergence judgment
$\NoRefine{\JJ}{\tau}$ expresses the inapplicability of $\tau$ to $\JJ$. In
practice, we will explain only the assertion conditions for one of
$\Refine{\JJ}{\tau}{\Psi}{E}$ and $\NoRefine{\JJ}{\tau}$, and implicitly take
the other to be its complement.

@@

In order to properly treat generic judgments whose syntheses bind variables, we
will commit up front to using the abstract binding trees logical framework as a
universal domain of discourse, with a distinguished sort $\SortJdg$ for
object-judgments; and to each judgment $\JJ$ such that
$\IsWf{\Omega}{\Upsilon}{\Gamma}{\JJ}{\SortJdg}$, a \emph{valence}
$\JdgValence{\JJ}$ shall be associated which describes the binding structure of
its synthesis. Then, to each subgoal context $\Psi$, we shall associate a
metavariable context $\Metavars{\Psi}$ as follows:

\begin{align*}
  \IMode{\Metavars{\cdot}} &= \OMode{\cdot}\\
  \IMode{\Metavars{\Psi, \alpha:\JJ}} &= \OMode{\Metavars{\Psi}, \alpha : \JdgValence{\JJ}}
\end{align*}

Then, the judgments $\NoRefine{\JJ}{\tau}$ and $\Refine{\JJ}{\tau}{\Psi}{E}$
shall presuppose $\IsWf{\cdot}{\cdot}{\cdot}{\JJ}{\SortJdg}$, and the latter
shall postsuppose $\IsWf{\Metavars{\Psi}}{\cdot}{\cdot}{E}{\JdgValence{\JJ}}$.
When $E$ binds no variables or symbols, we shall write $E$ instead of
$\MkBTm{\cdot}{\cdot}{E}$.

%% In practice, this means that the
%% judgment $\Refine{\JJ}{\tau}{\Psi}{E}$ shall presuppose
%% $\IsWf{\cdot}{\cdot}{\cdot}{\JJ}{\SortJdg}$

Let $\JSet$ be the open-ended collection of judgments in our logical theory,
and let $\RSet$ be a collection of rule names. Each rule $\Member{R}{\RSet}$
must be interpretable as a tactic, i.e.\ the meaning of the assertions
$\Refine{\JJ}{R}{\Psi}{E}$ and $\NoRefine{\JJ}{R}$ must be explained for
$\Member{J}{\JSet}$. We say \framebox{$\IsTac{\tau}$} in case for all
object-judgments $\Member{\JJ}{\JSet}$, the assertion conditions for
$\Refine{\JJ}{\tau}{\Psi}{E}$ and $\NoRefine{\JJ}{\tau}$ are well-defined and
disjoint.

@@

\begin{thm}
  The above rules all wellformed definitions of tactics.
  \begin{enumerate}
    \item $\IsTac{\IdTac}$.
    \item $\IsTac{\FailTac}$.
    \item If $\IsTac{\tau_1}$ and $\IsTac{\tau_2}$, then
      $\IsTac{\OrElseTac{\tau_1}{\tau_2}}$.

    \item If $\IsTac{\tau_1}$ and for any $\Member{\alpha}{I}$,
      $\IsTac{\tau_\alpha}$, then
      $\IsTac{\ThenTac{\tau_1}{\MkFam{\tau}{\alpha}{I}}}$.
  \end{enumerate}
\end{thm}
\begin{proof}
  Disjointness for each tactical is immediately evident; it suffices, then, to
  verify the postsupposition
  $\IsWf{\Metavars{\Psi}}{\cdot}{\cdot}{E}{\JdgValence{\JJ}}$ in each case.
  \begin{enumerate}
    \item[(1--2)] Immediate.
    \item[(3)]
      By induction on derivations of
      $\Refine{\JJ}{\OrElseTac{\tau_1}{\tau_1}}{E}{\Psi}$.
      \begin{description}
        \item[Case $\Refine{\JJ}{\tau_1}{\Psi}{E}$.] Validity follows from the
          inductive hypothesis $\IsTac{\tau_1}$.
        \item[Case $\NoRefine{\JJ}{\tau_1}$.] Validity follows from the
          inductive hypothesis $\IsTac{\tau_2}$.
      \end{description}

    \item[(4)] It suffices to verify the cases where
      $\IsEquiv{\Dom\Phi}{I}$, since otherwise
      $\NoRefine{\JJ}{\ThenTac{\tau_1}{\MkFam{\tau}{\alpha}{I}}}$.
      We need to show that
      $\SubsetEq{\FV{\Subst{F}{\Dom\Phi}{E}}}{\Dom{\bigoplus_{\Dom\Phi}\Psi}}$.
      From our inductive hypotheses, it is evident that
      $\SubsetEq{\FV{E}}{\Dom\Phi}$ and that for each $\Member{\alpha}{\Phi}$,
      $\SubsetEq{\FV{F_\alpha}}{\Dom{\Psi_\alpha}}$; all of the free metavariables
      of the term got by substituting each $F_\alpha$ for
      $\Member{\alpha}{\Dom\Phi}$ clearly reside in one of the fibres of the
      family of contexts $\Psi$, and so they must comprise a subset of the
      union $\bigoplus_{\Dom\Phi}\Psi$ of $\Psi$'s fibres.

@@

%%% TODO: correct to use new metavariable kit!!
